### 2022-01-06 response-to-review.md

# Response to peer review, Data Science submission 706-1686

Response to  open peer review at <https://datasciencehub.net/paper/packaging-research-artefacts-ro-crate> of “Packaging research artefacts with RO-Crate” submitted to Data Science._

We thank the reviewers for their attentive and detailed comments, and we very much appreciate the time and care they have given that has helped us improve the article.

## Response to Reviewer 1

_Open Review was by Angelo A. <https://orcid.org/0000-0002-4763-3943>_

We thank Angelo A for the comprehensive comments that we respond to below.

### Summary of changes

The reviewer raised important points on ensuring RO-Crate targets its aim of simplicity. This prompted us to include an explicit description of our user requirement and feedback methodology, with references to specific sections.

We have reflected this discussion in the following places:

* A new [section 2.1 Development Methodology](https://stain.github.io/ro-crate-paper/#methodology) that explicitly describe this requirement/feedback methodology in the article
* In section [3 RO-Crate Tooling](https://stain.github.io/ro-crate-paper/#tooling), by clarifying that the role of the tools aim to make creating and managing RO-Crates simple.
* A new section [7 Future Work](https://stain.github.io/ro-crate-paper/#futurework) that outlines some new directions and plans for RO-Crate 1.2 and a future user study linked to EOSC.

In addition, the reviewer has helped us improve the article with:

1. Language/typography fixes in [section 2](https://stain.github.io/ro-crate-paper/#rocrate), in [Figure 1](https://stain.github.io/ro-crate-paper/#fig:conceptual), [section 4.4](https://stain.github.io/ro-crate-paper/#dmp)
2. Missing [figure 4](https://stain.github.io/ro-crate-paper/#fig:hdc) restored
3. Citations added in [section 3](https://stain.github.io/ro-crate-paper/#tooling), [section 5.3](https://stain.github.io/ro-crate-paper/#packaging-workflows)
4. Clarifications added [section 2.2.3](https://stain.github.io/ro-crate-paper/#contextualentities),
5. Consistent section numbers HTML/PDF, added more cross references

### Detailed response per comment

@@

> I encourage the authors to add section numbers. Without them, it becomes hard to even scroll the document and reach the right location.

We agree - we have added section numbers to both PDF and HTML.

> **Typos**:
> - In section RO-Crate. “abuot” -> “about”

Fixed in [section 2](https://stain.github.io/ro-crate-paper/#rocrate).

> - Figure 1 caption. “The RO is described within a ,”. The object of the sentence is missing.

Fixed in [Figure 1](https://stain.github.io/ro-crate-paper/#fig:conceptual) - this was a Markdown error.

> - In Data Entities are described using Contextual Entities section, you wrote _“person, but secondarily also a web page and their academic work.”_ I think the sentence is incomplete.

The sentence was indeed incomplete. We have changed it in [section 2.2.3](https://stain.github.io/ro-crate-paper/#contextualentities) to:

_For instance https://orcid.org/0000-0002-1825-0097 is primarily an identifier for a person, but secondarily it is also a Web page and a way to refer to their academic work._

> - In Table 1, the tool SCHeMa has the squared parenthesis signalling a citation, but there are question marks.

Fixed in [section 3](https://stain.github.io/ro-crate-paper/#tooling).

> - In Machine-actionable Data Management Plans section, there is a left curly bracket ({)

Fixed in [section 4.4.](https://stain.github.io/ro-crate-paper/#dmp)

> - In Institutional data repositories – Harvard Data Commons, the reference to figure 3 is missing

Fixed [section 4.5](https://stain.github.io/ro-crate-paper/#institutionalrepos) with reference to (now) [figure 4](https://stain.github.io/ro-crate-paper/#fig:hdc).

> - In Institutional data repositories – Harvard Data Commons (PDF version) figure 3 is missing. In the HTML version, the figure 3 is available.

We have fixed this inconsistency (caused by graphics file extensions), Harvard Data Commons is now consistently [figure 4](https://stain.github.io/ro-crate-paper/#fig:hdc).

> - In Packaging Workflows, there is `[cite?]` signalling a missing citation.

We have reworded the sentence in [section 5.3](https://stain.github.io/ro-crate-paper/#packaging-workflows) to further clarify, adding the missing citation:

_Indeed, in the absence of a specific requirement and an editorial policy to back it up – such as eliminating the reference limit – authors are effectively discouraged from properly and comprehensively citing software [[114](https://doi.org/10.1038/s41592-019-0350-x)]._

> While reading the paper, I was eager to get to the evaluation section. At some point, in the Ensuring simplicity section, you state _“One aim of RO-Crate is to be conceptually simple. This simplicity has been repeatedly checked and confirmed through a community review process.”_
>
> I was indeed curious to see if you had **run a survey** on what users thought about RO-Crate. Perhaps, a survey gathering feedback on the user experience.

For our early development methodology we created a pool of early adopter projects from a range of disciplines and groups that we refer to in the article in new section [2.1 Development methodology](https://stain.github.io/ro-crate-paper/#methodology).

We are now ready for the next stage, which is to widen user adoption coupled with user evaluation to confirm our claims of conceptual simplicity and developer friendliness. We have added [section 7 Future Work](https://stain.github.io/ro-crate-paper/#futurework) where we outline our plans for developing novice-level tutorials and training events with feedback through user studies.

> Also, the concept of simplicity is bound with the skills of the practitioner. To what extent, they found it simple?

We have expanded on this in new section [2.1 Development methodology](https://stain.github.io/ro-crate-paper/#methodology).

@@

> Also, I trust that RO-Crate is far from being complete (actually, you are in the process of releasing v1.2). What do you think needs to be improved in RO-Crate? What other new features do you plan to include? How much of this is shaped by the end-users?

We have outlined the next directions and recent new contacts in new section [7 Future Work](https://stain.github.io/ro-crate-paper/#futurework).

## Response to Reviewer 2

@@

_By Alessia Bardi <https://orcid.org/0000-0002-1112-1292>_

We thank Alessia Bardi for her deep dive into the article, and for thoroughly checking the RO-Crate and the formal logic statements.

### Summary of changes

The reviewer found some bugs in the associated RO-Crate which we have fixed. We appreciate all the language improvements, which prompted us to revise the article readability.

1. Language and typographical fixes in [section 2.1](https://stain.github.io/ro-crate-paper/#conceptual), [Figure 1](https://stain.github.io/ro-crate-paper/#fig:conceptual), [section 2.2.1](https://stain.github.io/ro-crate-paper/#linkeddata), [section 2.4.1](https://stain.github.io/ro-crate-paper/#people), [section 4](https://stain.github.io/ro-crate-paper/#inuse), [section 4.4](https://stain.github.io/ro-crate-paper/#dmp), [section 4.5](https://stain.github.io/ro-crate-paper/#institutionalrepos), [section 9.2](https://stain.github.io/ro-crate-paper/#minimal-ro-crate)
2. Language/grammar revised throughout to improve readability (thanks Alan R Williams <https://orcid.org/0000-0003-3156-2105>)
3. Clarifications added in [section 2.2](https://stain.github.io/ro-crate-paper/#conceptual), [section 2.2.3](https://stain.github.io/ro-crate-paper/#contextualentities), [Figure 2](https://stain.github.io/ro-crate-paper/#fig:uml), [section 5.2](https://stain.github.io/ro-crate-paper/#fair-digital-objects), [section 7](https://stain.github.io/ro-crate-paper/#futurework)
4. [Listing 1](https://stain.github.io/ro-crate-paper/#jsonld) modified to show data entities refencing other data entities
5. References added [section 2.4.1](https://stain.github.io/ro-crate-paper/#people), [section 3](https://stain.github.io/ro-crate-paper/#tooling), [section 5.3](https://stain.github.io/ro-crate-paper/#packaging-workflows)
6. Fixed `ContextualEntity` mapping in formal logic statements of [appendix A section 9.2](https://stain.github.io/ro-crate-paper/#formaldefinition)

For the benefit of readers of _this response_ we have also below included an example of using the production rules from appendix A.

### Detailed response per comment

> The most important remark is that I had not a good experience with the RO-Crate of this publication.

> I clicked on supplementary files 1128.html and went to <https://datasciencehub.net/system/files/ds-supplementary-706-1128.html#ro-crate-metadata.json>

> - `conformsTo` is missing

We apologise for this bad experience and recognise that we should have done the validation that she did on our behalf. Mea culpa - we must “eat our own dog food”!

This turned out to be a [bug in ro-crate-html-js](https://github.com/UTS-eResearch/ro-crate-html-js/issues/10) rendering, it expected an contextual entity to display the linked RO-Crate standard. We have added such an entity in order to address this issue, but would point out this is not a formal requirement on a JSON level - see <https://www.researchobject.org/ro-crate/1.1/root-data-entity.html>

> - the link to download returns 404

This relative link was unfortunately broken by the datasciencehub.net submission website, which, out of our control, renamed the uploaded files and broke the relative links. We have checked with the IOS Press publisher, which unfortunately replied they are not able to host JSON files as supplementary material. In the accepted version we have published the corresponding RO-Crate separately using [GitHub pages](https://w3id.org/ro/doi/10.5281/zenodo.5146227) (with w3id.org PID) and Zenodo (DOI).

> - the RO-Crate check says it is not a valid RO-Crate

This was an unfortunate oversight in that the ro-crate-html-js rendering had a [more stringent view](https://github.com/UTS-eResearch/ro-crate-html-js/issues/11) than the specification on the requirements of license entities. We have fixed the updated RO-Crate.

> The same happens for the version deposited on Zenodo (https://doi.org/10.5281/zenodo.5146227) linked in the abstract.

We have updated the Zenodo deposit with the corrected RO-Crate.

_> Section "Conceptual definition":_

> - Change ';' with ':' in _"Figure 1 shows the main conceptual elements involved in an RO-Crate;"_

Fixed in [section 2.1](https://stain.github.io/ro-crate-paper/#conceptual).

> - Figure 1: check the caption, there are missing words at _"The RO is described within a , "_

Fixed [Figure 1](https://stain.github.io/ro-crate-paper/#fig:conceptual), this was an unfortunate Markdown/LaTeX rendering error.

> - Conceptual definition does not clarify what's the difference between a generic Research Object (as introduced in the previous section) and a RO-Crate

We have added to [section 2.2](https://stain.github.io/ro-crate-paper/#conceptual):

_The conceptual notion of a Research Object [[11](https://stain.github.io/ro-crate-paper/#ref-sRYUCzCq)] is thus realized with the RO-Crate model and serialized using Linked Data constructs within the RO-Crate metadata file._

> Section "Linked data as a foundation"
>
> - Move the footnote about IRIs the first time you use the acronym, if you think the footnote is needed

Agreed, IRI footnote moved to first mention in [section 2.2.1](https://stain.github.io/ro-crate-paper/#linkeddata).

> Section "Data Entities are described using Contextual Entities"
>
> The example of orcid is not very clear to me. Why would I need to include a contextual entity that describes the person? For example, as shown in listing 1, we have:
> 
>    	"author": {"@id": "https://orcid.org/0000-0002-1825-0097"}
>
> I can understand the suggestion to add
> 
>    	{ "@id": "https://orcid.org/0000-0002-1825-0097",
>         "@type": "Person",
>         "name": "Josiah Carberry"
>     },
>
> for discoverability (and convenience) purposes, but following the linked data practices I could get the jsonld of the person with
>
>     curl --location --request GET 'https://orcid.org/0000-0002-1825-0097' --header 'Accept: application/ld+json'

We are aware of this capability, and we believe other readers experienced with LOD will have similar thoughts. However, this showcases an important philosophical difference between RO-Crate and traditional Linked Data practices. We’ve added to [section 2.2.3](https://stain.github.io/ro-crate-paper/#contextualentities):

_In RO-Crate, a referenced contextual entity (e.g. a person identified by ORCID) should always be described within the RO-Crate Metadata File with at least a _type_ and _name_, even where their PID might resolve to further Linked Data. This is so that clients are not required to follow every link for presentation purposes, for instance HTML rendering. Similarly any imported [extension terms](https://www.researchobject.org/ro-crate/1.1/appendix/jsonld.html#extending-ro-crate) would themselves also have a human-readable description in the case where their PID does not resolve to human-readable documentation._

> Figure 2 does not clearly show how the data entities can be linked to each other. Neither does listing 1.

The UML is a simplification of a Linked Data model with an Open World Assumption. We have modified the [Figure 2](https://stain.github.io/ro-crate-paper/#fig:uml) title to _“Simplified UML class diagram of RO-Crate**”_ **and expanded _“_contextual entities_ (which may include aggregated and non-aggregated data entities). ”_

We have modified [Listing 1](https://stain.github.io/ro-crate-paper/#jsonld) to showcase this possibility of data entities linking each others, and added _“In this example `about `of the CSV data entity reference the `ImageObject`, which then take the roles of both a data entity and contextual entity”_

> Section "People"
>
> - add links to all mentioned projects.

In [section 2.4.1](https://stain.github.io/ro-crate-paper/#people) we have added links and citations for the projects.

> - Change ';' with ':' in "A key set of stakeholders are developers;"

Fixed in [section 2.4.1](https://stain.github.io/ro-crate-paper/#people).

> - Table 1: check missing link for SCHeMa

Fixed in [section 3](https://stain.github.io/ro-crate-paper/#tooling).

> Section "Profiles of RO-Crate in use"
>
> - Cultural Heritages: remove the final 's'

Removed in [section 4](https://stain.github.io/ro-crate-paper/#inuse).

> Section "Machine-actionable Data Management Plans"
>
> Pending brace at _" RO-Crate {RDA maDMP Mapper"_

Fixed in [section 4.4](https://stain.github.io/ro-crate-paper/#dmp).

> Section 'Institutional data repositories – Harvard Data Commons"
> 
> - missing figure number at "in the metadata to allow their future reuse (Figure~)"

Fixed in [section 4.5](https://stain.github.io/ro-crate-paper/#institutionalrepos)

> Section "FAIR Digital Objects"

> - the link behind "As an “abstract protocol”" returns 404

The link was accidentally broken by figure handling code in the manuscript HTML build process. Rephrased text in [section 5.2](https://stain.github.io/ro-crate-paper/#fair-digital-objects) to:

_The Digital Object Interface Protocol [[@doip2.0](https://www.dona.net/sites/default/files/2018-11/DOIPv2Spec_1.pdf)] can be considered an "abstract protocol" of requirements, DOs could be implemented in multiple ways._

> Section "packaging workflows"
>
> - Missing citation at _"authors might be penalised for doing so [cite?] "_

We rephrased this in [section 5.3](https://stain.github.io/ro-crate-paper/#packaging-workflows) to clarify, and added the missing citation:

_Indeed, in the absence of a specific requirement and an editorial policy to back it up – such as eliminating the reference limit – authors are effectively discouraged from properly and comprehensively citing software [[116](https://doi.org/10.1038/s41592-019-0350-x)]._

> Section "Minimal RO-Crate"
> 
> - Align terms in the definition of minimal RO-crate (`published` vs `datePublished`)
> 
> - Check footnote 8

Fixed to `datePublished()` in [section 9.2](https://stain.github.io/ro-crate-paper/#minimal-ro-crate) to be consistent [with footnote](https://stain.github.io/ro-crate-paper/#fn11).

> Section "Mapping to RDF with schema.org"
>
> - The mapping for ContextualEntity(e) is missing

Thanks for spotting this. For any interested readers, below we explore the mapping following the production rules.

The old mapping in [appendix A](https://stain.github.io/ro-crate-paper/#formaldefinition) assumed there was an explicit type(e, c) for the contextual entity, e.g. if we have:

    E ≡ <http://example.com/>
    L ≡ <https://spdx.org/licenses/CC0-1.0>
    license(E, L)

Then following the 𝕃𝖗𝖉𝖋 production rules:

    license(E, L)
    ⇒ Relation(E, &lt;http://schema.org/license>, L) ∧ CreativeWork(L)

    CreativeWork(L)
    ⇒ ContextualEntity(L) ∧

    type(L, <http://schema.org/CreativeWork>)
    ⇒ Relation(L, <http://schema.org/CreativeWork>)
    ∧ Class(<http://schema.org/CreativeWork>)

    Relation(E, <http://schema.org/license>, L)
    ⇒ Triple(E, <http://schema.org/license>, L)
    ∧IRI(E) ∧ IRI(L)

    Relation(L,<http://www.w3.org/1999/02/22-rdf-syntax-ns#type>, <http://schema.org/CreativeWork>)
    ⇒ Triple(L,<http://www.w3.org/1999/02/22-rdf-syntax-ns#type>,<http://schema.org/CreativeWork>)
    ∧IRI(L) ∧ IRI(&lt;http://schema.org/CreativeWork>)

In this example, `ContextualEntity(L)` did not need to get expanded to 𝕃𝖗𝖉𝖋, as we already got an implied type `CreativeWork(L)` from `license(E, L)`.

Expanding back E and L we get the corresponding triples:

    Triple(<http://example.com/>, <http://schema.org/license>, <https://spdx.org/licenses/CC0-1.0>)

    Triple(<https://spdx.org/licenses/CC0-1.0>, <http://www.w3.org/1999/02/22-rdf-syntax-ns#type>, <http://schema.org/CreativeWork>)

However this expectation of an explicit `type(e, l)` was indeed not enforced in our original definition of 𝕃𝖗𝖔𝖈𝖗𝖆𝖙𝖊.

Considering the recommendation in updated [section 2.2.3](https://stain.github.io/ro-crate-paper/#contextualentities): _“a referenced contextual entity should always be described within the RO-Crate Metadata File with at least a type and name, even where their PID might resolve to further Linked Data_”, then such a requirement should be added also to the first order logic rules.

Investigating this, you also helped uncover a variable name typo in the [section 9.2](https://stain.github.io/ro-crate-paper/#minimal-ro-crate) Class declaration:

    ∀e Entity(e) ⇒  Class(e)

In the updated manuscript, the above has been replaced with:

    ∀e Entity(e) ⇒  type(e, c) ∧ Class(c)
   	∀e ContextualEntity(e) ⇒  name(e, n)

The updated mapping to 𝕃𝖗𝖉𝖋 adds this relatively harmless fallback, to ensure a triple is produced for untyped URI references:

    ContextualEntity(e) ⇒ type(e, &lt;http://schema.org/Thing>)

The mapping for any arbitrary `ContextualEntity(E)` will then proceed as in the license example.

### 2022-01-06 content/80.formal-definition.tex
typo again!
schema.org -> Schema.org (in text)

\begin{eqnarray*}
ROCrate(R)                                  & \models & Root(R) \land Mentions(R, R) \land hasPart(R, d) \land \\
                                            & & Mentions(R, d) \land DataEntity(d) \land \\
                                            & & Mentions(R, c) \land ContextualEntity(c) \\
\forall r \ Root(r)                         & \Rightarrow & Dataset(r) \land name(r, n) \land description(r, d) \land \\
                                            & &             datePublished(r, date) \land license(e, l) \\
\forall e \forall n \ name(e, n)            & \Rightarrow & Value(n) \\
\forall e \forall s \ description(e, s)     & \Rightarrow & Value(s) \\
\forall e \forall d \ datePublished(e, d)   & \Rightarrow & Value(d) \\
\forall e \forall l \ license(e, l)         & \Rightarrow & ContextualEntity(l) \\
DataEntity(e)                               & \equiv &      File(e) \oplus Dataset(e) \\
Entity(e)                                   & \equiv &      DataEntity(e) \lor ContextualEntity(e) \\
\forall e \ Entity(e)                       & \Rightarrow & type(e, c) \land Class(e) \\
\forall e \ ContextualEntity(e)             & \Rightarrow & name(e, n)  \\
Mentions(R, s)                              & \models &     Relation(s, p, e) \oplus Attribute(s,  p, l) \\
Relation(s, p, o)                           & \models &     Entity(s) \land Property(p) \land  Entity(o) \\
Attribute(s, p, v)                          & \models &     Entity(s) \land Property(p) \land Value(v) \\
Value(v)                                    & \equiv &      v \in \mathbb{R} \oplus v \in \mathbb{S}
\end{eqnarray*}

@@

In reality many additional attributes and contextual types from schema.org types like \url{http://schema.org/affiliation} and \url{http://schema.org/Organization} would be used to further describe the RO-Crate and its entities, but as these are optional (\textit{SHOULD} requirements) they do not form part of this formalization.

\subsection{Mapping to RDF with schema.org}

@@

Below follows a mapping from $\mathcal{L}_{rocrate}$ to $\mathcal{L}_{RDF}$ using schema.org as vocabulary:

@@

Note that in the JSON-LD serialization of RO-Crate, the expression of $Class$ and $Property$ is typically indirect: The JSON-LD \texttt{@context} maps to schema.org IRIs, which, when resolved as Linked Data, embed their formal definition as RDFa. Extensions may however include such term definitions directly in the RO-Crate.

@@

Combining the above predicates and schema.org mapping with rudimentary JSON templates, these forward-chaining production rules can output JSON-LD according to the RO-Crate 1.1 specification\footnote{
    \textbf{Limitations:} Contextual entities not related from the RO-Crate (e.g. using inverse relations to a data entity) would not be covered by the single direction $Mentions(R, s)$ production rule; see \href{https://github.com/ResearchObject/ro-crate/issues/122}{GitHub issue ResearchObject/ro-crate\#122}. The $datePublished(e, d)$ rule do not include syntax checks for the ISO 8601 datetime format. Compared with RO-Crate examples, this generated JSON-LD does not use a $@context$ as the IRIs are produced unshortened, a post-step could do JSON-LD Flattening with a versioned RO-Crate context. The \texttt{@type} expansion is included for clarity, even though this is also implied by the $type(e, t)$ expansion to $Relation(e, \texttt{xsd:type})$.
}:

@@

This exposes the first order logic domain of discourse of IRIs, with rational numbers and strings as their corresponding JSON-LD representation. These production rules first grow the graph of $R$ by adding a transitive rule – anything described in $R$ which is related to $o$, means that $o$ is also mentioned by the $ROCrate(R)$. For simplicity this rule is one-way; in theory the graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities, but these are proposed to be bound to the root data entity with schema.org relation \href{http://schema.org/mentions}.

>>>>

\begin{eqnarray*}
ROCrate(R)                                  & \models & Root(R) \land Mentions(R, R) \land hasPart(R, d) \land \\
                                            & & Mentions(R, d) \land DataEntity(d) \land \\
                                            & & Mentions(R, c) \land ContextualEntity(c) \\
\forall r \ Root(r)                         & \Rightarrow & Dataset(r) \land name(r, n) \land description(r, d) \land \\
                                            & &             datePublished(r, date) \land license(e, l) \\
\forall e \forall n \ name(e, n)            & \Rightarrow & Value(n) \\
\forall e \forall s \ description(e, s)     & \Rightarrow & Value(s) \\
\forall e \forall d \ datePublished(e, d)   & \Rightarrow & Value(d) \\
\forall e \forall l \ license(e, l)         & \Rightarrow & ContextualEntity(l) \\
DataEntity(e)                               & \equiv &      File(e) \oplus Dataset(e) \\
Entity(e)                                   & \equiv &      DataEntity(e) \lor ContextualEntity(e) \\
\forall e \ Entity(e)                       & \Rightarrow & type(e, c) \land Class(c) \\
\forall e \ ContextualEntity(e)             & \Rightarrow & name(e, n)  \\
Mentions(R, s)                              & \models &     Relation(s, p, e) \oplus Attribute(s,  p, l) \\
Relation(s, p, o)                           & \models &     Entity(s) \land Property(p) \land  Entity(o) \\
Attribute(s, p, v)                          & \models &     Entity(s) \land Property(p) \land Value(v) \\
Value(v)                                    & \equiv &      v \in \mathbb{R} \oplus v \in \mathbb{S}
\end{eqnarray*}

@@

In reality many additional attributes and contextual types from Schema.org types like \url{http://schema.org/affiliation} and \url{http://schema.org/Organization} would be used to further describe the RO-Crate and its entities, but as these are optional (\textit{SHOULD} requirements) they do not form part of this formalization.

\subsection{Mapping to RDF with Schema.org}

@@

Below follows a mapping from $\mathcal{L}_{rocrate}$ to $\mathcal{L}_{RDF}$ using Schema.org as vocabulary:

@@

Note that in the JSON-LD serialization of RO-Crate, the expression of $Class$ and $Property$ is typically indirect: The JSON-LD \texttt{@context} maps to Schema.org IRIs, which, when resolved as Linked Data, embed their formal definition as RDFa. Extensions may however include such term definitions directly in the RO-Crate.

@@

Combining the above predicates and Schema.org mapping with rudimentary JSON templates, these forward-chaining production rules can output JSON-LD according to the RO-Crate 1.1 specification\footnote{
    \textbf{Limitations:} Contextual entities not related from the RO-Crate (e.g. using inverse relations to a data entity) would not be covered by the single direction $Mentions(R, s)$ production rule; see \href{https://github.com/ResearchObject/ro-crate/issues/122}{GitHub issue ResearchObject/ro-crate\#122}. The $datePublished(e, d)$ rule do not include syntax checks for the ISO 8601 datetime format. Compared with RO-Crate examples, this generated JSON-LD does not use a $@context$ as the IRIs are produced unshortened, a post-step could do JSON-LD Flattening with a versioned RO-Crate context. The \texttt{@type} expansion is included for clarity, even though this is also implied by the $type(e, t)$ expansion to $Relation(e, \texttt{xsd:type})$.
}:

@@

This exposes the first order logic domain of discourse of IRIs, with rational numbers and strings as their corresponding JSON-LD representation. These production rules first grow the graph of $R$ by adding a transitive rule – anything described in $R$ which is related to $o$, means that $o$ is also mentioned by the $ROCrate(R)$. For simplicity this rule is one-way; in theory the graph can also contain free-standing contextual entities that have outgoing relations to data- and contextual entities, but these are proposed to be bound to the root data entity with Schema.org relation \href{http://schema.org/mentions}.
